IMF and Benchmark Forecasts

### 2023-10-17 manuscript/manuscript.tex

\begin{itemize}
\item Inflation: directional vs. absolute errors:
\begin{itemize}
\item difference small for IMF method, absolute slighly better, likely due to longer central intervals
\item AR and BVAR profit more from directional correction (upward bias)
\item for expanding window method, difference in coverage is smaller (-> structural breaks)

@@

\begin{verbatim}
## Warning: Using alpha for a discrete variable is not advised.
## Using alpha for a discrete variable is not advised.
\end{verbatim}

\begin{verbatim}
## Warning: Removed 1 rows containing missing values (`geom_bar()`).
\end{verbatim}

\includegraphics{manuscript_files/figure-latex/gdpovr_ew-1.pdf}

@@

\includegraphics{manuscript_files/figure-latex/overall_assessment-1.pdf}
\newpage

@@

\begin{verbatim}
## Warning: Using alpha for a discrete variable is not advised.
## Using alpha for a discrete variable is not advised.
\end{verbatim}

\begin{verbatim}
## Warning: Removed 1 rows containing missing values (`geom_bar()`).
\end{verbatim}

>>>>

\begin{itemize}
\item Inflation: directional vs. absolute errors:
\begin{itemize}
\item difference small for IMF method, absolute slightly better, likely due to longer central intervals
\item AR and BVAR profit more from directional correction (upward bias)
\item for expanding window method, difference in coverage is smaller (-> structural breaks)

@@

\includegraphics{manuscript_files/figure-latex/gdpovr_rw-1.pdf}

@@

\includegraphics{manuscript_files/figure-latex/infklovr_ew-1.pdf}

@@

### 2023-10-17 manuscript/sub_manuscripts/quantile_coverage.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\usepackage{lineno}
\linenumbers
\newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
\usepackage{float}
\usepackage{booktabs}
\usepackage{caption}
\usepackage{longtable}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={IMF and Benchmark Forecasts},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

@@

\setstretch{2}
\hypertarget{extracting-error-quantiles}{%
\section{Extracting error quantiles}\label{extracting-error-quantiles}}

Consider a forecast that stems from a source \(s\) for a specific target \(k\) in a country \(j\), for target year \(t\) and with forecast horizon \(h\):
\[
\hat{y}_{s,k,j,t,h}
\]
For example, this could be a forecast stemming from the International Monetary Fund World Economic Outlook (\(s = IMF\)) for real GDP growth (\(k = gdpg\)) in Canada (\(j = Canada\)) for the year \(2022\) (\(t=2022\)). \(h\) then indexes the forecast horizon, where we code:
\[
h = \left\{
\begin{array}{ll} 
      0, & \textrm{for forecasts made in October of the same year} \\
      0.5, & \textrm{for forecasts made in April of the same year} \\
      1, & \textrm{for forecasts made in October of the previous year} \\
      1.5, & \textrm{for forecasts made in April of the previous year} \\
\end{array}\right. 
\]
After the target year has completed, we obtain the realized value for the quantity of interest. For these, the WEO updates publishes biannual updates for two years, yielding \(4\) versions of the realized value. In accordance with previous literature (\textit{cite Timmermann 2008}), we use the version that is published in October of the following year and thereby don't index the true value by its publishing date (\textit{rephrase}). We thus write the true value as
\[
\hat{y}_{k,j,t}
\]
Given the forecast and the realized value for the quantity of interest, we can calculate the respective forecast error as
\[
e^{d}_{s,k,j,t,h} = y_{k,j,t} - \hat{y}_{s,k,j,t,h}
\]

for the ``directional'' error method and as
\[
e^{a}_{s,k,j,t,h} = \left|y_{k,j,t} - \hat{y}_{s,k,j,t,h}\right|
\]
for the ``absolute'' error method.

The objective is to extract quantiles from sets of errors \(\mathcal{E}_{s,k,j,t,h}\) constructed of certain years, depending on the estimation method \(m\), to be able to quantify the uncertainty inherent in the forecasts via central prediction intervals of level \(\alpha = \{0.5, 0.8\}\). For the estimation method, we consider a ``rolling window'' method, an ``expanding window'' method, and a ``leave-one-out'' method. For the rolling window method (\(m = rw\)), the errors of the last nine years enter into the estimation. For the expanding window method (\(m = ew\)), all previous years are considered, leaving a nine year window up front for the first estimation. For the leave-one-out method, all years except the current target year enter the estimation set. The latter is of course equivalent to the expanding window method in a real time setting and is considered in the scope of this analysis as a mere check \textit{rephrase}.
As an example, the error set for the ``directional'' error method and the rolling window approach is
\[
\mathcal{E}^{d,rw}_{s,k,j,t,h} =  \left\{  e^{d}_{s,k,j,t^*,h} | t-9\leq t^* < t \right\}
\]
Insert reasoning to use the past \(9\) errors.

To now obtain the lower \(l\) and upper \(u\) values for a central prediction interval of level \(\alpha\), we take quantiles of these sets and add them to the current prediction:

For the directional method:
\[
l^{\alpha, d}_{t,h,v,l,j} = \hat{y}_{t, h, l, j} + q^{0.5 - \alpha/2 } \left(\mathcal{E}^{d, m}_{t,h,v,l,j}  \right)
\]
\[
u^{\alpha, d}_{t,h,v,l,j} = \hat{y}_{t, h, l, j} + q^{0.5 + \alpha/2 } \left(\mathcal{E}^{d, m}_{t,h,v,l,j}  \right)
\]

And for the absolute method:

\[
 l^{\alpha, a}_{t,h,v,l,j} = \hat{y}_{t, h, l, j} -  q^{\alpha} \left(\mathcal{E}^{m, a}_{t,h,v,l,j}  \right)
\]
\[
u^{\alpha, a}_{t,h,v,l,j} = \hat{y}_{t, h, l, j} +  q^{\alpha} \left(\mathcal{E}^{m, a}_{t,h,v,l,j}  \right)
\]

Two different philosophies.

The absolute method will always yield symmetric central prediction intervals around the forecast value, while the directional method will in general yield asymmetric intervals. They thus result in different central intervals, unless the errors in \(\mathcal{E}\) are perfectly symmetric around zero\footnote{Not totally correct, actually. For this to hold exactly, the error set would need to be augmented with one zero value.}. In fact, the directional method can yield central prediction intervals that do not even contain the forecast value, in cases where the \((0.5 - \alpha/2)\)-quantile is positive or the \((0.5 + \alpha/2)\)-quantile is negative.

\hypertarget{scores-by-error-method-horizon-and-forecast-source}{%
\section{Scores, by error method, Horizon and forecast source}\label{scores-by-error-method-horizon-and-forecast-source}}

\hypertarget{inflation}{%
\subsection{Inflation}\label{inflation}}

\includegraphics{manuscript_files/figure-latex/overall_assessment_ew-1.pdf}
\newpage
Some notes:

\begin{itemize}
\item Inflation: directional vs. absolute errors:
\begin{itemize}
\item difference small for IMF method, absolute slightly better, likely due to longer central intervals
\item AR and BVAR profit more from directional correction (upward bias)
\item for expanding window method, difference in coverage is smaller (-> structural breaks)

\end{itemize}
\item Inflation overall scores: IMF forecasts outperform others
\begin{itemize}
\item lower scores for point forecasts
\item lower WIS 
\item lower bias (compute directly?)
\end{itemize}

\item GDP Growth: more similar results for different sources
\begin{itemize}
\item lower scores at shorter horizons, more similar at larger horizons
\item IMF forecasts better only for absolute error method
\end{itemize}
\end{itemize}

\newpage

\hypertarget{gdp}{%
\subsection{GDP}\label{gdp}}

\includegraphics{manuscript_files/figure-latex/gdpovr_rw-1.pdf}

\hypertarget{expanding-window---scores-by-error-method-horizon-and-forecast-source}{%
\section{Expanding Window - Scores, by error method, Horizon and forecast source}\label{expanding-window---scores-by-error-method-horizon-and-forecast-source}}

\hypertarget{inflation-1}{%
\subsection{Inflation}\label{inflation-1}}

\includegraphics{manuscript_files/figure-latex/infklovr_ew-1.pdf}

\newpage

\hypertarget{gdp-1}{%
\subsection{GDP}\label{gdp-1}}

\includegraphics{manuscript_files/figure-latex/gdpovr-1.pdf}

\hypertarget{coverage-by-target-methods-and-source}{%
\section{Coverage, by target, methods and source}\label{coverage-by-target-methods-and-source}}

\includegraphics{manuscript_files/figure-latex/cvgplot-1.pdf}

\newpage
\begin{longtable}{l|rrr}
\toprule
\multicolumn{1}{l}{} & IMF & ar & bvar \\ 
\midrule
\multicolumn{4}{l}{horizon = 0} \\ 
\midrule
expanding window\_interval\_score & 0.115 & 0.123 & 0.122 \\ 
expanding window\_sample\_crps & 0.087 & 0.092 & 0.091 \\ 
leave-one-out\_interval\_score & 0.112 & 0.123 & 0.123 \\ 
leave-one-out\_sample\_crps & 0.088 & 0.099 & 0.100 \\ 
rolling window\_interval\_score & 0.119 & 0.128 & 0.126 \\ 
rolling window\_sample\_crps & 0.094 & 0.097 & 0.098 \\ 
\midrule
\multicolumn{4}{l}{horizon = 0.5} \\ 
\midrule
expanding window\_interval\_score & 0.258 & 0.272 & 0.296 \\ 
expanding window\_sample\_crps & 0.182 & 0.208 & 0.241 \\ 
leave-one-out\_interval\_score & 0.257 & 0.257 & 0.276 \\ 
leave-one-out\_sample\_crps & 0.180 & 0.217 & 0.230 \\ 
rolling window\_interval\_score & 0.261 & 0.266 & 0.288 \\ 
rolling window\_sample\_crps & 0.191 & 0.201 & 0.235 \\ 
\midrule
\multicolumn{4}{l}{horizon = 1} \\ 
\midrule
expanding window\_interval\_score & 0.448 & 0.737 & 0.590 \\ 
expanding window\_sample\_crps & 0.327 & 0.504 & 0.426 \\ 
leave-one-out\_interval\_score & 0.443 & 0.746 & 0.587 \\ 
leave-one-out\_sample\_crps & 0.302 & 0.448 & 0.400 \\ 
rolling window\_interval\_score & 0.451 & 0.739 & 0.591 \\ 
rolling window\_sample\_crps & 0.333 & 0.514 & 0.434 \\ 
\midrule
\multicolumn{4}{l}{horizon = 1.5} \\ 
\midrule
expanding window\_interval\_score & 0.495 & 1.044 & 0.800 \\ 
expanding window\_sample\_crps & 0.346 & 0.627 & 0.602 \\ 
leave-one-out\_interval\_score & 0.488 & 1.021 & 0.756 \\ 
leave-one-out\_sample\_crps & 0.337 & 0.583 & 0.577 \\ 
rolling window\_interval\_score & 0.494 & 1.041 & 0.779 \\ 
rolling window\_sample\_crps & 0.347 & 0.632 & 0.570 \\ 
\bottomrule
\end{longtable}

\hypertarget{inflation-2}{%
\subsection{Inflation}\label{inflation-2}}

\includegraphics{manuscript_files/figure-latex/cpitileplot-1.pdf}

\hypertarget{gdp-2}{%
\subsection{GDP}\label{gdp-2}}

\begin{longtable}{l|rrr}
\toprule
\multicolumn{1}{l}{} & IMF & ar & bvar \\ 
\midrule
\multicolumn{4}{l}{horizon = 0} \\ 
\midrule
expanding window\_interval\_score & 0.241 & 0.301 & 0.298 \\ 
expanding window\_sample\_crps & 0.178 & 0.209 & 0.204 \\ 
leave-one-out\_interval\_score & 0.235 & 0.297 & 0.295 \\ 
leave-one-out\_sample\_crps & 0.183 & 0.219 & 0.214 \\ 
rolling window\_interval\_score & 0.240 & 0.302 & 0.297 \\ 
rolling window\_sample\_crps & 0.178 & 0.211 & 0.204 \\ 
\midrule
\multicolumn{4}{l}{horizon = 0.5} \\ 
\midrule
expanding window\_interval\_score & 0.416 & 0.540 & 0.493 \\ 
expanding window\_sample\_crps & 0.298 & 0.383 & 0.358 \\ 
leave-one-out\_interval\_score & 0.406 & 0.532 & 0.482 \\ 
leave-one-out\_sample\_crps & 0.310 & 0.408 & 0.377 \\ 
rolling window\_interval\_score & 0.416 & 0.554 & 0.504 \\ 
rolling window\_sample\_crps & 0.297 & 0.405 & 0.373 \\ 
\midrule
\multicolumn{4}{l}{horizon = 1} \\ 
\midrule
expanding window\_interval\_score & 0.837 & 1.090 & 0.942 \\ 
expanding window\_sample\_crps & 0.640 & 0.822 & 0.763 \\ 
leave-one-out\_interval\_score & 0.833 & 1.083 & 0.933 \\ 
leave-one-out\_sample\_crps & 0.634 & 0.759 & 0.709 \\ 
rolling window\_interval\_score & 0.851 & 1.122 & 0.970 \\ 
rolling window\_sample\_crps & 0.663 & 0.875 & 0.815 \\ 
\midrule
\multicolumn{4}{l}{horizon = 1.5} \\ 
\midrule
expanding window\_interval\_score & 1.045 & 1.288 & 1.102 \\ 
expanding window\_sample\_crps & 0.790 & 0.937 & 0.897 \\ 
leave-one-out\_interval\_score & 1.030 & 1.292 & 1.109 \\ 
leave-one-out\_sample\_crps & 0.765 & 0.882 & 0.849 \\ 
rolling window\_interval\_score & 1.056 & 1.312 & 1.143 \\ 
rolling window\_sample\_crps & 0.831 & 0.985 & 0.949 \\ 
\bottomrule
\end{longtable}

\includegraphics{manuscript_files/figure-latex/gdptileplot-1.pdf}

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\usepackage{lineno}
\linenumbers
\newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
\usepackage{float}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={IMF and Benchmark Forecasts},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

@@

\setstretch{2}
\hypertarget{this-document}{%
\section{This document}\label{this-document}}

Here we document the proportion of true values that ``miss'' the 50\% and 80\% central prediction intervals - that is, the proportions of observations that falls \textit{below} the 0.1-quantile and 0.25-quantile and the proportion that falls \textit{above} the 0.75-quantile and 0.9-quantile.

The respective proportions are calculated separately for each forecast source (IMF, ar, bvar, bvarqu\footnote{bvarqu denotes the quantile forecasts that are taken directly from the BVAR model, while bvar denotes the quantile forecasts that are drawn from past point forecast errors of the BVAR model (analagously to the IMF and ar forecasts)}, target (GDP Growth and Inflation) and horizon (0 - ``Fall, same year'' to 1.5 - ``Spring, previous year''). We treat absolute error handling as the default, and show results for directional error handling in the same respective source's color, but in a lighter shade. This plot is shown on page 3, with desired nominal exceedance proportions shown as dashed lines.

On page 4, we also plot exceedance levels for the bvarqu model separately for two equally sized parts of the time period under study. Differences in coverage between the two time periods at the respective quantile levels seem non-substantial.

\hypertarget{gdp-growth-forecasts}{%
\section{GDP growth forecasts}\label{gdp-growth-forecasts}}

Overall, the forecasts that stem (a) from the directional error methods for the ar and bvar forecasts and (b) the IMF forecasts in general tend to be better calibrated, especially at the lower quantile levels.

The bvarqu forecasts show relatively large\footnote{in relation to the respective desired nominal level} exceedance proportions at the lower quantile levels (q = 0.1, q = 0.25) and small exceedance proportions at the upper quantile levels, an indication of overprediction. The respective forecast intervals ``cover too much ground'' in the direction of positive deviations of the realized values from the point prediction, while simultaneously ``covering too little'' in the other direction. Similarly, this is also an issue for the absolute ar, bvar and (albeit to a slightly lesser extent) IMF forecasts.

Lastly, the bvarqu forecasts appear to be too narrow at the shortest horizon (0 - ``Fall, same year'') only, since they show relatively large exceedances at both the lower and the upper quantile levels.

\hypertarget{inflation-forecasts}{%
\section{Inflation forecasts}\label{inflation-forecasts}}

For the inflation forecasts, the bvarqu model (along with absolute ar and bvar to a lesser extent) show small exceedance proportions at the upper quantiles, especially for the 0.9-quantile and larger horizons, indicating intervals that are too long in the region above the point forecast. This also explains the aggregate near-zero underprediction scores for these methods (see other pdf).

Interestingly, the directional methods show the opposite issue, in that they have a too large proportion of exceedances at the upper quantile levels.

At the lower quantile levels, all methods appear to be better calibrated.

\hypertarget{overall}{%
\section{Overall}\label{overall}}

In summary, it appears that for GDP, the entire distributions seem to be biased (shifted) upwards for the bvarqu and absolute ar and bvar models. For inflation, it seems that they are just too long at the top, while being better calibrated at the bottom.

\includegraphics{quantile_coverage_files/figure-latex/unnamed-chunk-4-1.pdf}

\hypertarget{exceedance-for-bvarqu-by-time-period}{%
\section{Exceedance for bvarqu, by time period}\label{exceedance-for-bvarqu-by-time-period}}

\includegraphics{quantile_coverage_files/figure-latex/unnamed-chunk-5-1.pdf}
